---
title: "BM2_HW2"
author: "Yangyang Chen"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r, include = FALSE, message = FALSE, warning = FALSE}
library(tidyverse)
library(viridis)
library(modelr)
library(mgcv)
library(latex2exp)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

## Problem 1

### (a) Fill out the table and give comments.

```{r}
# load data
dose=c(0:4)## x_i
num=rep(30, 5) ## m_i
killed=c(2, 8, 15, 23, 27) ## y_i
data=data.frame(dose,num,killed) ## (x_i, m_i, y_i)

# data preparation
x=data$dose
y=data$killed
m=data$num
resp=cbind(y,m-y)   #### counts of success (death=1), failure (surv=0)

# Model fitting
# logit link
glm_logit=glm(resp~x, family=binomial(link='logit'))
# probit link
glm_probit=glm(resp~x, family=binomial(link='probit'))
# complementary log-log link
glm_clog=glm(resp~x, family=binomial(link='cloglog')) # asymmetric

results_func = function(fit,model){
  fit_summary = fit |> summary()
  beta = fit_summary$coefficients[2]
  CI = fit$coefficients +
    kronecker(t(c(0,qnorm(0.025),-qnorm(0.025))), t(t(sqrt(diag(vcov(fit))))))
  CI_lower = CI[2,2]
  CI_upper = CI[2,3]
  dev = fit_summary$deviance
  p_new = predict.glm(fit, newdata = tibble(x = 0.01), type = "response")
  return(tibble(model, beta, CI_lower, CI_upper, dev, p_new))
}

results_func(glm_logit, "logit") |> 
  rbind(results_func(glm_probit, "probit")) |> 
  rbind(results_func(glm_clog, "c-log-log")) |> 
  knitr::kable(digits = 4)
```

Comments:

* All models showed a significant relationship between dose and death with significant p-values less than 0.001.

* As the dose increases, the probability of death decreases because all CIs for the dose estimates are positive and do not include zero.

* The probit link model fits the data best since it minimizes residual deviance.

* The probability of death estimated using the probit link model at a dose level of 0.01 was 0.0853.

### (b) Suppose that the dose level is in natural logarithm scale, estimate LD50 with 90% confidence interval based on the three models.

Given:

$$Var(\hat{x_0}) = (\frac{\partial{x_0}}{\partial{\alpha}})^2Var(\hat{\alpha}) + (\frac{\partial{x_0}}{\partial{\beta}})^2Var(\hat{\beta}) +
2(\frac{\partial{x_0}}{\partial{\alpha}})(\frac{\partial{x_0}}{\partial{\beta}})Cov(\hat{\alpha},\hat{\beta})$$

90% CI: $$[\hat{x_0} - Z_{0.95}*\sqrt{Var(\hat{x_0})}, \hat{x_0} + Z_{0.95}*\sqrt{Var(\hat{x_0})}]$$.

Point estimate for LD50 is $exp(\hat{x_0})$; 90% CI for LD50 is $[e^{\hat{X_L}}, e^{\hat{X_R}}]$.

For model using logit link, when $\pi_0 = 0.5,\ g(\pi_0) = log(\frac{\pi_0}{1-\pi{0}}) = 0 = \alpha + \beta X$. For model using probit link, when $\pi_0 = 0.5, \ g(\pi_0) = {\phi}^{-1}(\pi_0) = 0 = \alpha + \beta X$. For model using comlimentary log-log link, when $\pi_0 = 0.5, g(\pi_0) = log(-log(1-\pi_0)) = log(-log(0.5)) = \alpha + \beta X$.

Then, for the models using lgit link and probit link,

$$\hat{x_0} = - \frac{\hat{\alpha}}{\hat{\beta}}$$

$$var(\hat{x_0}) =
\frac{1}{\hat{\beta^2}}Var(\hat{\alpha})+
\frac{\hat{\alpha^2}}{\hat{\beta}^4}Var(\hat{\beta})-
2\frac{\hat{\alpha}}{\hat{\beta^3}}Cov(\hat{\alpha},\hat{\beta})$$

If we use complementary log-log link,

$$\hat{x_0} = \frac{log(-log(0.5))-\hat{\alpha}}{\hat{\beta}}$$

$$var(\hat{x_0}) = \frac{1}{\hat{\beta^2}}Var(\hat{\alpha})+
\frac{(log(-log(0.5))-\hat{\alpha})^2}{\hat{\beta}^4}Var(\hat{\beta})+
2\frac{log(-log(0.5))-\hat{\alpha}}{\hat{\beta^3}}Cov(\hat{\alpha},\hat{\beta})$$

Implementation:
```{r}
LD50_func = function(fit, model){
  alpha = fit$coefficients[1]
  beta = fit$coefficients[2]
  betacov = vcov(fit)
  if(model == "c-log-log"){
    x0fit = (log(-log(0.5)) - alpha) /beta
    varx0 = betacov[1,1]/(beta^2)+
      betacov[2,2]*(log(-log(0.5)) - alpha) ^ 2 / (beta ^ 4) + 2 * betacov[1,2] * (log(-log(0.5)) - alpha) / (beta ^ 3)
  }
  else
  {
    x0fit = -alpha / beta
    varx0 = betacov[1,1] / (beta ^ 2) +
      betacov[2,2] * (alpha ^ 2) / (beta ^ 4) - 
      2 * betacov[1,2] *alpha / (beta ^ 3)
  }
  estimate = exp(x0fit)
  CI_lower = exp(x0fit - c(qnorm(0.95)) * sqrt(varx0))
  CI_upper = exp(x0fit + c(qnorm(0.95))*sqrt(varx0))
  return(tibble(model, estimate, CI_lower, CI_upper))
}

LD50_func(glm_logit, "logit") |> 
  rbind(LD50_func(glm_probit, "probit")) |> 
  rbind(LD50_func(glm_clog, "c-log-log")) |> 
  knitr::kable(digits = 2)
```

## Problem 2

Use logistics regression with logit link, we obtained model:

$$log odds(Enrolls) = \hat{\alpha} + \hat{\beta} * Amount$$
```{r}
amount=seq(10, 90, by = 5)
offers=c(4, 6, 10, 12, 39, 36, 22, 14, 10, 12, 8, 9, 3, 1, 5, 2, 1)
enrolls=c(0, 2, 4, 2, 12, 14, 10, 7, 5, 5, 3, 5, 2, 0, 4, 2, 1)
data=data.frame(amount, offers, enrolls)
```

```{r}
glm_logit_2 = glm(cbind(enrolls, offers - enrolls) ~ amount,
                family = binomial(link = "logit"))
glm_logit_2 |> summary()
```

### (a) How does the model fit the data?

Since most of the offers are less than 30, we can assume that the data is sparse. Therefore, we should use the Hosmer-Lemeshow statistic to evaluate the goodness of fit:

```{r}
library(ResourceSelection)
```

```{r}
hoslem_stat = hoslem.test(glm_logit_2$y, fitted(glm_logit_2), g = 10)
hoslem_stat
```

* Since $p-value = 0.9907 > 0.05$, we failed to reject the null hypothesis and therefore concluded that the model fits the data well at 95% significant level.

### (b) How do you interpret the relationship between the scholarship amount and enrollment rate? What is 95% CI?

```{r}
logit_summary = glm_logit_2 |> summary()
beta_exp = exp(logit_summary$coefficients[2])
CI_exp = exp(glm_logit_2$coefficients + 
               kronecker(t(c(0,qnorm(0.025), -qnorm(0.025))),t(t(sqrt(diag(vcov(glm_logit_2)))))))
CI_beta_exp = str_c("[", CI_exp[2,2] |> round(2), ", ", CI_exp[2,3] |> round(2), "]", sep = "")
```

* The model shows that there is a significant relationship between the scholarship amount and enrollment rate. 

* The 95% confidence interval is [1.01, 1.05].

### (c) How much scholarship should we provide to get 40% yield rate (the percentage of admitted students who enroll?) What is the 95% CI?

When $\pi_0 = 0.4, g(\pi_0) = log(\frac {\pi_0}{1-\pi_0}) = log(\frac{2}{3}) = \alpha + \beta X$ Then,
$$\hat{x_0} = \frac{log(\frac{2}{3})-\hat{\alpha}}{\hat{\beta}}$$

$$var(\hat{x_0}) = \frac{1}{\hat{\beta^2}}Var(\hat{\alpha})+\frac{log(\frac{2}{3})-\hat{\alpha}}{\hat{\beta}^4}Var(\hat{\beta})+2\frac{log(\frac{2}{3})-\hat{\alpha}}{\hat{\beta^3}}Cov(\hat{\alpha},\hat{\beta})$$

Therefore, we can estimate the scholarship amount and its 95% CI as follows:

```{r}
alpha_hat = glm_logit_2$coefficients[1]
beta_hat = glm_logit_2$coefficients[2]
beta_cov = vcov(glm_logit_2) # inverse of fisher-information
x0_hat = (log(2/3) - alpha_hat) / beta_hat
var_x0 = 
  beta_cov[1,1] / (beta_hat^2) +
 (log(2/3) - alpha_hat)^2 / (beta_hat^4) * beta_cov[2,2] +
  2 * (log(2/3) - alpha_hat) / beta_hat^3 * beta_cov[1,2]
CI_lower = x0_hat - c(qnorm(0.975)) * sqrt(var_x0)
CI_upper = x0_hat + c(qnorm(0.975)) * sqrt(var_x0)
tibble(estimate = x0_hat, CI_lower, CI_upper) |> knitr::kable(digits = 4)
```

* Therefore, approximately 40,000 dollars scholarship should be provided to get 40% yield rate.

* The 95% CI is [30.583, 49.686].


